\begin{center}
    \textsc{Acknowledgments}
  \end{center}
  \noindent

First and foremost, this dissertation is a result of many supportive advisors, teachers, colleagues, family and friends - who accompanied me through my studies and showed me the values to pursue in both research and personal life.

I would like to show my deepest gratitude to my supervisors Dr. Ahmed Bouajjani and Dr. Constantin Enea. Without their assistance and advice in every step throughout the process, even in the time of pandemic, this thesis would have not been accomplished. They taught me many valuable lessons about research, to be a good researcher and to always challenge myself. I would like to thank you very much for your support and understanding over the last three and half years.

Also I would like to show my gratitude to Dr. Bengt Jonsson (Professor, Uppsala University, Sweden), Dr. Ilya Sergey (Associate Professor, Yale-NUS College, Singapore), Dr. Mihaela Sighireanu (Professor, ENS Paris-Saclay, France), Dr. Noam Rinetzky (Associate Professor, Tel Aviv University, Israel), Dr. Viktor Vafeiadis (Tenured Researcher, Max Planck Institute for Software Systems (MPI-SWS), Kaiserslautern) for agreeing to be the jury of my dissertation. Bengt and Ilya were very kind to review my dissertation in this challenging time and accommodate the delay in the completion of its writing. Thank you all very much. 

I also like to thank my colleagues and friends Sidi Mohamed Beillahi, Berk Cirisci, Suha Orhun Mutluergil, Rachid Zennou. Whenever I was stuck at some problems, the discussions with you helped me to progress. Thanks for providing a friendly environment at IRIF. I would also like to thank the administrative staff at ED 386, IRIF, and UFR Informatique for their help during my stay at IRIF and my dissertation.

I also like to thank my faculties in undergrad university, Chennai Mathematical Institute (Chennai, India). Thank you Mr. Balaguru Srivathsan, for your lectures that inspired me to choose Computer Science as my career. This PhD opportunity would have never been possible without your guidance and advice.

Most importantly, none of this could have ever happened without my family. My mother Kanyakumarika Biswas, my father Ramtanu Biswas and my grandparents Namita Paik and Krishnapada Paik and my brothers Rudradeep Biswas and Suvadeep Biswas and my uncle Arup Paik. Without your love, sacrifice and emotional and financial support, I would never reach the place where I am right now.

But last but not least, this journey would not have been interesting and fun without my friends. My school friends  Poulomi Aich, Mouli Bose, Sanchari Chakraborty, Subhajit Mukherjee and Kanchan Yadav - thanks for your continuous support and friendship. My university friends Sougata Bose, Abhishek De, Arpita Kar - thanks for all the career advice and friendship. Darshana Krishnamony - thank you for being the first helping hand and friend in Paris and staying that way.  Anjum Mullani - thank you for adopting a brother and teaching me to be a better person in life. You are the sister I never had. Melisa Benakay and Simone Perriello - thank you for keeping me sane with all the best experiences in Europe. You are the best of friends that I could never imagine. Thank you Paula Souto Montero for all the love and support. You have been an amazing and loving person.

At the end, I would like to thank each and everyone in my life, who contributed to make me a better person and reach this position. Thank you.